---
title: Exploratory Data Analysis (EDA)
subtitle: ""
# author:
# - affiliation: Instituto Agroforestal Mediterráneo. Departamento de Ecosistemas Agroforestales.
#     Universitat Politècnica de València.
#   email: algarsal@upv.es
#   name: Alfonso Garmendia
title-block-banner: true
date: last-modified
date-format: iso
lang: en-EN
language: 
  title-block-published: "Last modified"
code-fold: true
format:
  html:
    toc: true
    toc-depth: 6
    toc-location: body
    toc-title: Table of contents
    embed-resources: true
    self-contained-math: true
---

```{r setup, include=FALSE}
library(skimr)

```

## Import data

```{r import_data}
# dir("data")
load("data/clean-data.Rdata")
```

Data are imported from the file *"data/clean-data.Rdata"*. It was created running the script "*01_data-survey-clean-export.qmd*" and has several dataframes: *`r ls()`*. 

See the document [01_data-survey-clean-export.html](01_data-survey-clean-export.html) for a detailed description of the data creation and legend. 

## Exploratory Data Analysis (EDA)

EDA is important to check the structure of the data frame and basic correlations. 

There are some automatic EDA tools [here](https://finnstats.com/index.php/2021/05/04/exploratory-data-analysis/) and [here](https://www.business-science.io/code-tools/2023/01/05/exploratory-dataxray-analysis.html).

## EDA with `skimr` package
### EDA for *pathog* dataframe

```{r}
skimr::skim(pathog)
```

### EDA for *nguy* dataframe

```{r}
skimr::skim(nguy)
```

### EDA for *samu* dataframe

```{r}
skimr::skim(samu)
```


## Correlation funnel

In the figures for correlation funnel we can see the variables ordered from the most correlated to the appearance of symptoms and which values are more with or without them. 

```{r}

library(correlationfunnel)
library(ggplot2)
# names(pathog)
v <- c("species", "inoc_part", "batch",
  "symptom", "days", "host", "variety",
  "sample", "location", "province")
pathog$days <- as.numeric(pathog$days)
pathog[v] %>% 
  binarize() %>% 
  correlate(symptom__0) %>% 
  plot_correlation_funnel() + 
  labs(title = "Correlation Funnel for pathog dataframe")

nguy$days <- as.numeric(nguy$days)
nguy[v] %>% 
  binarize() %>% 
  correlate(symptom__0) %>% 
  plot_correlation_funnel() + 
  labs(title = "Correlation Funnel for nguy dataframe")

samu$days <- as.numeric(samu$days)
samu[v] %>% 
  binarize() %>% 
  correlate(symptom__0) %>% 
  plot_correlation_funnel() + 
  labs(title = "Correlation Funnel for samu dataframe")


```


# {{< pagebreak >}}

```{r}
#| eval: !expr file.exists("rmd/ch_reproducibility.qmd")
#| child: "rmd/ch_reproducibility.qmd"

```




<!-- ## Bibliography -->
