---
title: Pathogeneity models
subtitle: "samuen"
# author:
# - affiliation: Instituto Agroforestal Mediterráneo. Departamento de Ecosistemas Agroforestales.
#     Universitat Politècnica de València.
#   email: algarsal@upv.es
#   name: Alfonso Garmendia
title-block-banner: true
date: last-modified
date-format: iso
lang: en-EN
language: 
  title-block-published: "Last modified"
code-fold: true
format:
  html:
    toc: true
    toc-depth: 6
    toc-location: body
    toc-title: Table of contents
    embed-resources: true
    self-contained-math: true
---

```{r setup, include=FALSE}
library(knitr)
library(agricolae)
```

## Import data

```{r import_data}
# dir("data")
load("data/clean-data.Rdata")
rm(pathog, nguy)

### By species --------
carob <- samu[samu$species == "CAROB", ]
manda <- samu[samu$species ==
    "CLEMENVILLA.MANDARIN", ]
persi <- samu[samu$species == "PERSIMMON", ]

```

Data are imported from the file *"data/clean-data.Rdata"*. It was created running the script "*01_data-survey-clean-export.qmd*" and has several dataframes: *`r ls()`*. 

See the document [01_data-survey-clean-export.html](01_data-survey-clean-export.html) for a detailed description of the data creation and legend. 

See the document [02_EDA](02_EDA.html) for a quick survey of the data. 

### Strains and species

```{r, eval=FALSE}
table(samu$code, samu$species)
```

## Pathogenicity of strains

¿Which strains are pathogenic? 

To test this we order the levels of *code* and do an anova and Tukey post-hoc test but only comparing the Control (without strains) with each strain. 

### On carob leaves

```{r}
### Order levels of code to put "CONTROL" 
### or "none" as the first level -----------
# levels(carob$location)
carob$code <- relevel(carob$code, "CONTROL")

### Anova ------------------
asc <- aov(symptom ~ code, data = carob)
tasc0 <- TukeyHSD(asc)$code
tasc1 <- tasc0[grep("CONTROL", row.names(tasc0)), ]

atasc0 <- HSD.test(asc, "code")

### KW -----------------
# kw0 <- agricolae::kruskal(carob$symptom, carob$code)

### Shapiro test residuals -----------
sh0 <- shapiro.test(asc$residuals)
```

```{r tbl-tasc1}
#| tbl-cap: Differences between control and each strain. 

kable(tasc1)

```

```{r fig-tasc0}
#| fig-cap: Mean pathogenicity of each strain. Different letters for significance (alpha = 0.05).

plot(atasc0, las = 2)

# plot(kw0, las = 2)

```


### On mandarine leaves

```{r}
### Order levels of code to put "CONTROL" 
### or "none" as the first level -----------
# levels(manda$location)
manda$code <- relevel(manda$code, "CONTROL")

### Anova ------------------
asc <- aov(symptom ~ code, data = manda)
tasc0 <- TukeyHSD(asc)$code
tasc1 <- tasc0[grep("CONTROL", row.names(tasc0)), ]

atasc0 <- HSD.test(asc, "code")

### KW -----------------
# kw0 <- agricolae::kruskal(manda$symptom, manda$code)

### Shapiro test residuals -----------
sh0 <- shapiro.test(asc$residuals)
```

```{r tbl-tasc1-m}
#| tbl-cap: Differences between control and each strain. 

kable(tasc1)

```

```{r fig-tasc0-m}
#| fig-cap: Mean pathogenicity of each strain. Different letters for significance (alpha = 0.05).

plot(atasc0, las = 2)

# plot(kw0, las = 2)

```

### On persimon leaves

```{r}
### Order levels of code to put "CONTROL" 
### or "none" as the first level -----------
# levels(persi$location)
persi$code <- relevel(persi$code, "CONTROL")

### Anova ------------------
asc <- aov(symptom ~ code, data = persi)
tasc0 <- TukeyHSD(asc)$code
tasc1 <- tasc0[grep("CONTROL", row.names(tasc0)), ]

atasc0 <- HSD.test(asc, "code")

### KW -----------------
# kw0 <- agricolae::kruskal(persi$symptom, persi$code)

### Shapiro test residuals -----------
sh0 <- shapiro.test(asc$residuals)
```

```{r tbl-tasc1-p}
#| tbl-cap: Differences between control and each strain. 

kable(tasc1)

```

```{r fig-tasc0-p}
#| fig-cap: Mean pathogenicity of each strain. Different letters for significance (alpha = 0.05).

plot(atasc0, las = 2)

# plot(kw0, las = 2)

```


# {{< pagebreak >}}

```{r}
#| eval: !expr file.exists("rmd/ch_reproducibility.qmd")
#| child: "rmd/ch_reproducibility.qmd"

```




<!-- ## Bibliography -->
