---
title: Model diagram
subtitle: "Samu"
# author:
# - affiliation: Instituto Agroforestal Mediterráneo. Departamento de Ecosistemas Agroforestales.
#     Universitat Politècnica de València.
#   email: algarsal@upv.es
#   name: Alfonso Garmendia
title-block-banner: true
date: last-modified
date-format: iso
lang: en-EN
language: 
  title-block-published: "Last modified"
code-fold: true
format:
  html:
    toc: true
    toc-depth: 6
    toc-location: body
    toc-title: Table of contents
    embed-resources: true
    self-contained-math: true
---

```{r import_data, eval=FALSE}
# dir("data")
load("data/clean-data.Rdata")
rm(pathog, nguy)
```


## Variables

The independent variable is the strain (code) and the dependent variable is symptoms. 

Leaves (inoc_part: young - old) or days (from inoculation to evaluation) are variables independent to strain but can affect symptoms. 

Species is a confounder, because it affects both symptoms and the strains used. 

The strain characteristics can be moderators, affected by strain and affecting symptoms. 

Batch should not have an effect. 

See @fig-simple 

```{mermaid}
%%| label: fig-simple
%%| fig-cap: "Possible relationships between variables."

flowchart LR
  St --> Sc(Strain characteristics: \n host, \n variety, \n sample, \n location, \n province.)
  Sc --> Sy
  St{strain} ==> Sy{symptoms}
  Sp(species) --> Sy
  Sp(species) --> St
  Le(leaves) --> Sy 
  Da(Days) --> Sy
  Ba(batch) -..-> Sy
```

\

# {{< pagebreak >}}

```{r}
#| eval: !expr file.exists("rmd/ch_reproducibility.qmd")
#| child: "rmd/ch_reproducibility.qmd"

```




<!-- ## Bibliography -->
