---
title: Plots
subtitle: "Nguyen"
# author:
# - affiliation: Instituto Agroforestal Mediterráneo. Departamento de Ecosistemas Agroforestales.
#     Universitat Politècnica de València.
#   email: algarsal@upv.es
#   name: Alfonso Garmendia
title-block-banner: true
date: last-modified
date-format: iso
lang: en-EN
language: 
  title-block-published: "Last modified"
code-fold: true
format:
  html:
    toc: true
    toc-depth: 6
    toc-location: body
    toc-title: Table of contents
    embed-resources: true
    self-contained-math: true
---

```{r setup, include=FALSE}
library(ggplot2)
library(dplyr)

```

## Data

```{r import_data}
# dir("data")
load("data/clean-data.Rdata")
rm(pathog, samu)
```

Data are imported from the file *"data/clean-data.Rdata"*. It was created running the script "*01_data-survey-clean-export.qmd*" and has several dataframes: *`r ls()`*. 

See the document [01_data-survey-clean-export.html](01_data-survey-clean-export.html) for a detailed description of the data creation and legend. 

See the document [02_EDA](02_EDA.html) for a quick survey of the data. 

In the document [04n_model-diagram-Nguyen.html](04n_model-diagram-Nguyen.html) there is a diagram of the model possible relationships to test. 

```{r}
### Order levels of code to 
### put "CONTROL" or "none" as the first level
# levels(nguy$location)
nguy$code <- relevel(nguy$code, "CONTROL")

```

All factor variables were reordered so control observations are in the first place. 

## Best model

```{r}
call <- "symptom ~ code * inoc_part"
```

After comparing a set of different models, the best fitted model is (see [05n_models-comparison-Nguyen.html](05n_models-comparison-Nguyen.html)):

- `r call`

For all strains but AAPVC5 the effect in young leaves is stronger than in adult leaves. The effect is different for each strain (see @fig-ssi).

```{r fig-ssi}
#| fig-cap: Effect of the strain and the age of the inoculated leaves (inoculated_in) on the Alternaria disease symptoms. Error bars stand for standard deviation. Asterics denote significant differences with the control experiment (alpha = 0.05). 
#| fig-width: 9
#| column: screen

### Analysis -------------
aad <- aov(symptom ~ code, 
  data = nguy[nguy$inoc_part == "ADULT LEAVES", ])
ayo <- aov(symptom ~ code, 
  data = nguy[nguy$inoc_part == "YOUNG LEAVES", ])

aadt <- as.data.frame(TukeyHSD(aad)$code)
aadt <- aadt[grep("CONTROL", row.names(aadt)), ]
aadt$inoc_part <- "ADULT LEAVES"
aadt$sig <- ""
aadt$sig[aadt$`p adj` < 0.05] <- "*"

ayot <- as.data.frame(TukeyHSD(ayo)$code)
ayot <- ayot[grep("CONTROL", row.names(ayot)), ]
ayot$inoc_part <- "YOUNG LEAVES"
ayot$sig <- ""
ayot$sig[ayot$`p adj` < 0.05] <- "*"
scontrol <- aadt[1, ]
scontrol$sig <- ""
s <- rbind(scontrol, aadt)
s <- rbind(s, scontrol)
s <- rbind(s, ayot)
s$code <- rep(levels(nguy$code), 2)
sig <- s[c("code", "inoc_part", "sig")]

### Data for plot -------------
d <- nguy %>%
  group_by(code, inoc_part) %>%
  summarise_at(vars(symptom), 
    list(mean = mean, sd = sd)) %>% 
  as.data.frame()

### plot ----------------------
pd <- position_dodge(0.5)
ggplot(d, aes(x = code, y = mean, 
  color = inoc_part)) + 
  geom_errorbar(
    aes(ymin = mean - sd, ymax = mean + sd), 
    width = .3, position = pd) +
  geom_point(size = 2, position = pd) +
  geom_text(data = sig, position = pd,
    size = 8,
    aes(y = 4.5, x = code,
      color = inoc_part, label = sig)) +
  labs(x = "Strain", y = "symptom level mean", 
    color = "inoculated in") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, 
    vjust = 1, hjust = 1))  


```

It could be interesting to replicate the graph for variety, sample, location and province to see if it is easier to understand. 


# {{< pagebreak >}}

```{r}
#| eval: !expr file.exists("rmd/ch_reproducibility.qmd")
#| child: "rmd/ch_reproducibility.qmd"

```




<!-- ## Bibliography -->
