---
title: Plots
subtitle: "Samu"
# author:
# - affiliation: Instituto Agroforestal Mediterráneo. Departamento de Ecosistemas Agroforestales.
#     Universitat Politècnica de València.
#   email: algarsal@upv.es
#   name: Alfonso Garmendia
title-block-banner: true
date: last-modified
date-format: iso
lang: en-EN
language: 
  title-block-published: "Last modified"
code-fold: true
format:
  html:
    toc: true
    toc-depth: 6
    toc-location: body
    toc-title: Table of contents
    embed-resources: true
    self-contained-math: true
---

```{r setup, include=FALSE}
library(ggplot2)
library(dplyr)

```

## Data

```{r import_data}
# dir("data")
load("data/clean-data.Rdata")
rm(pathog, nguy)
```

Data are imported from the file *"data/clean-data.Rdata"*. It was created running the script "*01_data-survey-clean-export.qmd*" and has several dataframes: *`r ls()`*. 

See the document [01_data-survey-clean-export.html](01_data-survey-clean-export.html) for a detailed description of the data creation and legend. 

See the document [02_EDA](02_EDA.html) for a quick survey of the data. 

In the document [04s_model-diagram-Samu.html](04s_model-diagram-Samu.html) there is a diagram of the model possible relationships to test. 

```{r}
### Order levels of code to 
### put "CONTROL" or "none" as the first level
# levels(samu$location)
samu$host <- relevel(samu$host, "none")
samu$code <- relevel(samu$code, "CONTROL")
```

All factor variables were reordered so control observations are in the first place. 

## Best model

```{r}
call <- "symptom ~ species * host * province"
```

After comparing a set of different models, the best fitted model is (see [05s_models-comparison-Samu.html](05s_models-comparison-Samu.html)):

- `r call`


```{r fig-ssi}
#| fig-cap: Effect of the species, host and province on the Alternaria disease symptoms. Error bars stand for standard deviation. Asterics denote significant differences with the control experiment (alpha = 0.05). 
#| fig-width: 9
#| column: screen

### Data for plot -------------
d <- samu %>%
  group_by(species, host) %>%
  summarise_at(vars(symptom), 
    list(mean = mean, sd = sd)) %>% 
  as.data.frame()

### plot ----------------------
pd <- position_dodge(0.5)
ggplot(d, aes(x = host, y = mean, 
  color = species)) + 
  geom_errorbar(
    aes(ymin = mean - sd, ymax = mean + sd), 
    width = .3, position = pd) +
  geom_point(size = 2, position = pd) +
  # geom_text(data = sig, position = pd,
  #   size = 8,
  #   aes(y = 4.5, x = code,
  #     color = inoc_part, label = sig)) +
  labs(x = "host strain", y = "symptom level mean", 
    color = "infected species") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 10, 
    vjust = 1, hjust = 1))  


```

Experiment for Persimmon cannot be used because the control (not inoculated) was  with very strong symptoms. This species had strong symptoms in all experiments. It is very sensible to *Alternaria*. 

For the other strains, only the strains obtained from carob tree did cause symptoms in either carob tree and mandarine. 

The effect of carob tree strains on carob tree leaves is always 2, without variation (see @fig-carob).

The effect of carob tree strains in carob tree leaves is always the same two and clearly different from the control.

The effect of carob tree strains in mandarin leaves depends on the strain (see @fig-mand).

```{r}
tf <- samu$species == "CAROB" & 
  samu$host %in% c("none", "CAROB TREE")
d <- samu[tf, ]
d$symptomTF <- d$symptom > 0
d$host <- droplevels(d$host)
### analysis --------
t1 <- table(d$host, d$symptomTF)
ch <- fisher.test(t1)

```


```{r fig-carob}
#| fig-cap: !expr paste("Frequencies of symptoms in carob tree leaves for the strains from carob tree, compared with the control. p.value of fisher test is:", format(ch$p.value, digits = 2), ".")

#### Data 
D <- as.data.frame(t1)
D$total <-
  # sums all same names in Var1
  sapply(D$Var1,
    function(caseID)
      sum(subset(D,
        Var1 == caseID)$Freq))
D$percent <- round((D$Freq / D$total) * 100, 2)
D <- D[order(D$Var2, decreasing = TRUE),]
D <- D[order(D$Var1),]

### plot  -------------
  ### First plot ####
  bps <- ggplot(D, aes(x = Var1, y = Freq, fill = Var2)) +
    geom_bar(stat = "identity", position = "stack") +
    labs(x = "carob", y = "Frequency")
  
  ### Frequencies ####
  freq = ggplot_build(bps)$data[[1]]
  freq$y_pos = (freq$ymin + freq$ymax) / 2
  Dord <- D[order(D$Var2, decreasing = TRUE),]
  Dord <- Dord[order(Dord$Var1),]
  
  bps <- bps +
    annotate(
      x = freq$x,
      y = freq$y_pos,
      label = D$Freq,
      geom = "text",
      size = 5,
      color = "black"
    )
  
  #### Legend and theme ####
  bps <- bps +
    guides(fill = guide_legend(title = "Symptoms"), color = "none") 
bps
```

```{r fig-mand}
#| fig-cap: Effect of carob strains on mandarine leaves. Error bars stand for standard deviation. Asterics denote significant differences with the control experiment (alpha = 0.05). 
#| fig-width: 9
#| column: screen

### Data for plot -------------
tf <- samu$species == "CLEMENVILLA.MANDARIN" & 
  samu$host %in% c("none", "CAROB TREE")

d0 <- samu[tf, ] 
d0$code <- droplevels(d0$code)
d <- d0 %>%
  group_by(inoc_part, code) %>%
  summarise_at(vars(symptom), 
    list(mean = mean, sd = sd)) %>% 
  as.data.frame()

### Analysis -------------
aad <- aov(symptom ~ code, 
  data = d0[d0$inoc_part == "ADULT LEAVES", ])
ayo <- aov(symptom ~ code, 
  data = d0[d0$inoc_part == "YOUNG LEAVES", ])

aadt <- as.data.frame(TukeyHSD(aad)$code)
aadt <- aadt[grep("CONTROL", row.names(aadt)), ]
aadt$inoc_part <- "ADULT LEAVES"
aadt$sig <- ""
aadt$sig[aadt$`p adj` < 0.05] <- "*"

ayot <- as.data.frame(TukeyHSD(ayo)$code)
ayot <- ayot[grep("CONTROL", row.names(ayot)), ]
ayot$inoc_part <- "YOUNG LEAVES"
ayot$sig <- ""
ayot$sig[ayot$`p adj` < 0.05] <- "*"
scontrol <- aadt[1, ]
scontrol$sig <- ""
s <- rbind(scontrol, aadt)
s <- rbind(s, scontrol)
s <- rbind(s, ayot)
s$code <- rep(levels(d0$code), 2)
sig <- s[c("code", "inoc_part", "sig")]

### plot ----------------------
pd <- position_dodge(0.5)
ggplot(d, aes(x = code, y = mean, 
  color = inoc_part)) + 
  geom_errorbar(
    aes(ymin = mean - sd, ymax = mean + sd), 
    width = .3, position = pd) +
  geom_point(size = 2, position = pd) +
  geom_text(data = sig, position = pd,
    size = 8,
    aes(y = 4.5, x = code,
      color = inoc_part, label = sig)) +
  labs(x = "code", y = "symptom level mean", 
    color = "inoculated leaves") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 10, 
    vjust = 1, hjust = 1))  


```

# {{< pagebreak >}}

```{r}
#| eval: !expr file.exists("rmd/ch_reproducibility.qmd")
#| child: "rmd/ch_reproducibility.qmd"

```




<!-- ## Bibliography -->
