---
title: "Cross Pathogenicty on Carob, Clemenvilla and Persimmon"
author: "Mohamed Samu"
date: "2023-06-07"
output: html_document 
---

```{r}
library(tidyverse)
library(openxlsx)
library(ggplot2)
library(writexl)
library(readxl)
library(agricolae)
library(xtable)
library(knitr)
library(kableExtra)
library(xlsx)
library(PMCMRplus)
```

## Load the data

Three previously isolated pathogenic Alternaria a. strains of Mandarin, Carob, and Persimmon from different locations were used in this study.

```{r}
load("data/clean-data.Rdata")

# samu
# pathog

```

##Table 1: Description of Alternaria spp. used in the study. \*\*CODE: Identification code of the strain; *HOST: Common name of the host species;* VARIETY: Name of the variety; *LOCATION: Municipality collected from;* PROVINCE Province of the municipality; and \*ORCHARD: Name of the orchard\*\*

```{r}
Table1 <- pathog[, c(1, 3, 10, 11, 13, 14)]

excel_file <- "Table1.xlsx"  # Specify the desired file name and path

write.xlsx(Table1, file = excel_file, row.names = FALSE)


```

\***The result comprise analysis for 33 strain for which, some show no symptom. Analyzing all the data will give major statistical problem. So, we first summarize the the result and create a table showing the mean and standard deviation for all the strains tested.**

##Table 2: Result sumarry of mean and standard deviation **Results covers all the 33 Alternarian a. strains and the crops on which they were tested.**

```{r}
#Summarize the result and create a table 
Result0 <- samu%>% group_by(code, species)%>%
  summarise(Mean=mean(symptom),
            sd=sd(symptom),
            max= max(symptom), 
            min=min(symptom))    
```

##Analysis For pathogenicity on CLEMENVILLA.MANDARIN

*Select, Analyse and plot for only those showing pathogenic on CLEMENVILLA.MANDARIN and control*

```{r}
##Select and Analyse for only infected strain of CLEMENVILLA.MANDARIN and control
df1 <- samu%>% filter(species=="CLEMENVILLA.MANDARIN")
df <- df1[df1$code %in% c("CAROB_S1", "CAROB_S4", "CAROB_S5", "CAROB1", "CONTROL"), ]
a<- kruskal(df$symptom, df$code, console = TRUE,group = TRUE)
letter <- a$groups$groups

df2 <- df%>% group_by(species, code)%>%
  summarise(Mean=mean(symptom),
            sd=sd(symptom),
            max= max(symptom), 
            min=min(symptom))      
 
ggplot(df2)+geom_bar(stat="identity", aes(y=Mean, x=code, color =species), width=0.5, position = "dodge")+ 
theme_bw()+ggtitle("Cross_Pathogenicity on Young_Leaves")+
  labs(x = "host strain", y ="Mean Severity Index", color = "infected species")+
   geom_text(
     aes(x =1:5, y = Mean, label = letter),
     position = position_dodge(width = 0.5),
     vjust = -0.5, size = 3, hjust = 2,
   )+
theme_bw() +
  theme(axis.text.x = element_text(angle = 10))


#### Add an Error bar to the plot
ggplot(df2) +
  geom_bar(stat = "identity", aes(y = Mean, x = code, fill = species), width = 0.5, position = "dodge") +
  geom_errorbar(aes(ymin = Mean - sd, ymax = Mean + sd, x = code), width = 0.5, position = position_dodge(width = 0.5)) +  # Error bars
  scale_fill_manual(values = c("red", "black")) +  # Set species colors to green and brown
  theme_bw() +
  ggtitle("Cross Pathogenicity on Young Leaves") +
  labs(x = "host strain", y = "Mean Severity Index", fill = "infected species") +
  geom_text(
    aes(x = 1:5, y = Mean + sd, label = letter),
    position = position_dodge(width = 0.5),
    vjust = -0.5, size = 5, hjust = 0.5, color = "black"
  ) +
  theme(axis.text.x = element_text(angle = 10))

```

*Analyse and plot for all the strains tested on CLEMENVILLA.MANDARIN*

```{r}
##Select and Analyse for all of all the strains CLEMENVILLA.MANDARIN
df3 <- samu%>% filter(species=="CLEMENVILLA.MANDARIN")
b<- kruskal(df3$symptom, df3$code, console = TRUE,group = TRUE)
letter0 <- b$groups$groups

df4 <- df3%>% group_by(species, code)%>%
  summarise(Mean=mean(symptom),
            sd=sd(symptom),
            max= max(symptom), 
            min=min(symptom))      
 
ggplot(df4)+geom_bar(stat="identity", aes(y=Mean, x=code, color =species), width=0.5, 
                      position = "dodge")+
theme_bw()+ggtitle("Cross_Pathogenicity on Young_Leaves")+
  labs(x = "Host Strain", y ="Mean Severity Index", color = "Infected Species")+
   geom_text(
     aes(x =1:9, y = Mean + sd, label = letter0),
     position = position_dodge(width = 0.001),
     vjust = -0.5, size = 3, hjust = 2,
   )+
theme_bw() +
  theme(axis.text.x = element_text(angle = 10))

#### Add an Error bar to the plot
ggplot(df4) +
  geom_bar(stat = "identity", aes(y = Mean, x = code, fill = species), width = 0.5, position = "dodge") +
  geom_errorbar(aes(ymin = Mean - sd, ymax = Mean + sd, x = code), width = 0.5, position = position_dodge(width = 0.5)) +
  scale_fill_manual(values = c("red", "black")) +
  theme_bw() +
  ggtitle("Cross Pathogenicity on Young Leaves") +
  labs(x = "Host Strain", y = "Mean Severity Index", fill = "Infected Species") +
  geom_text(
    aes(y = Mean + sd, x = code, label = letter0),
    position = position_dodge(width = 0.5),
    vjust = -0.5, size = 5, hjust = 0.5, color = "black"
  ) +
  theme(axis.text.x = element_text(angle = 10))

```

##Analysis For pathogenicity on PERSIMMON

*For Persimmon, all the strains tested including control showered pathogenicity*

```{r}
##Select and Analyse for only infected strain of PERSIMMON and control
Per <- samu%>% filter(species=="PERSIMMON")
P<- kruskal(Per$symptom, Per$code, console = TRUE,group = TRUE)
Plet <- P$groups$groups

Per0 <- Per%>% group_by(species, code)%>%
  summarise(Mean=mean(symptom),
            sd=sd(symptom),
            max= max(symptom), 
            min=min(symptom))      

ggplot(Per0)+geom_bar(stat="identity", aes(y=Mean, x=code, color =species), width=0.5, 
                      position = "dodge")+
theme_bw()+ggtitle("Cross_Pathogenicity on Young_Leaves")+
  labs(x = "Host Strain", y ="Mean Severity Index", color = "Infected Species")+
   geom_text(
     aes(x =1:12, y = Mean, label = Plet),
     position = position_dodge(width = 0.001),
     vjust = -0.5, size = 3, hjust = 2,
   )+
theme_bw() +
  theme(axis.text.x = element_text(angle = 10))

#### Add an Error bar to the plot
ggplot(Per0) +
  geom_bar(stat = "identity", aes(y = Mean, x = code, fill = species), width = 0.5, position = "dodge") +
  geom_errorbar(aes(ymin = Mean - sd, ymax = Mean + sd, x = code), width = 0.2, position = position_dodge(width = 0.1)) +  # Error bars
  scale_fill_manual(values = c("blue", "black")) +  # Set species colors to green and brown
  theme_bw() +
  ggtitle("Cross_Pathogenicity on Young_Leaves") +
  labs(x = "Host Strain", y = "Mean Severity Index", fill = "Infected Species") +
  geom_text(
    aes(x = 1:12, y = Mean + sd, label = Plet),
    position = position_dodge(width = 0.5),
    vjust = -0.5, size = 5, hjust = 0.5, color = "black"
  ) +
  theme(axis.text.x = element_text(angle = 10))

```

##Analysis For pathogenicity on CAROB

*For Persimmon, all the strains tested including control showered pathogenicity*

```{r}
library(agricolae)
##Select and Analyse for only infected strain of PERSIMMON and control
carob <- samu%>% filter(species=="CAROB")
car<- kruskal(carob$symptom, carob$code, console = TRUE,group = TRUE)
Clet <- result$groups$groups

library(coin)

# Perform Kruskal-Wallis test with ties
result <- kruskal_test(Car$symptom ~ Car$code)

# Perform post hoc multiple comparisons
posthoc <- posthoc.kruskal.nemenyi.test(Car$symptom, Car$code)

# Get the letters of significance
n <- letters(posthoc$p.value < 0.05)

# View the letters of significance
print(letters)

Car0 <- Car%>% group_by(species, code)%>%
  summarise(Mean=mean(symptom),
            sd=sd(symptom),
            max= max(symptom), 
            min=min(symptom))    

```
