---
title: "Cross Pathogenicty on Carob, Clemenvilla and Persimmon"
author: "Mohamed Samu"
date: "2023-06-07"
output: html_document 
---

```{r}
library(tidyverse)
library(openxlsx)
library(ggplot2)
library(writexl)
library(readxl)
library(agricolae)
library(xtable)
```

## Load the data
Three previously isolated pathogenic Alternaria a. strains of Mandarin, Carob, and Persimmon from different locations were used in this study.

```{r}
load("data/clean-data.Rdata")

# samu
# pathog

```

##Table 1: Description of Alternaria spp. used in the study. 
**CODE:	Identification code of the strain; *HOST:	Common name of the host species; *VARIETY:	Name of the variety; *LOCATION:	Municipality collected from; *PROVINCE	Province of the municipality; and *ORCHARD:	Name of the orchard**

```{r}
Table1 <- pathog[, c(1, 3, 10, 11, 13, 14)]

```

***The result comprise analysis for 33 strain for which, some show no symptom. Analyzing all the data will give major statistical problem. So, we first summarize the the result and create a table showing the mean and standard deviation for all the strains tested.** 

##Table 2: Result sumarry of mean and standard deviation
**Results covers all the 33 Alternarian a. strains and the crops on which they were tested.**  
```{r}
#Summarize the result and create a table 
Result0 <- samu%>% group_by(code, species)%>%
  summarise(Mean=mean(symptom),
            sd=sd(symptom),
            max= max(symptom), 
            min=min(symptom))    
```

##Analysis by inoculum for only pathogenic trains 

Result sumarry of mean and standard deviation

```{r}
##Select and Analyse for CLEMENVILLA.MANDARIN
df1 <- samu%>% filter(species=="CLEMENVILLA.MANDARIN")
df <- df1[df1$code %in% c("CAROB_S1", "CAROB_S4", "CAROB_S5", "CAROB1", "CONTROL"), ]
a<- kruskal(df$symptom, df$code, console = TRUE,group = TRUE)
letter <- a$groups$groups

df2 <- df%>% group_by(species, code)%>%
  summarise(Mean=mean(symptom),
            sd=sd(symptom),
            max= max(symptom), 
            min=min(symptom))      
 
ggplot(df2)+geom_bar(stat="identity", aes(y=Mean, x=code, fill=species), width=0.5, 
                      position = "dodge")+
theme_bw()+ggtitle("Young_leaves")+xlab("Mean Severity Index")+ylab("Pathogen Strains")+
   geom_text(
     aes(x =1:5, y = Mean, label = letter),
     position = position_dodge(width = 1),
     vjust = -0.5, size = 5
   )

```

